Long-form recording of infant body position in the home using wearable inertial sensors

### 2023-06-14 manuscript.Rmd
some revision edits

From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings. Changes in *body position*---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction. When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position [@Freeplay; @LuoFranchak2020; @CWW]. Infants struggle to manipulate objects while supine and prone, but sitting affords object exploration [@SoskaAdolph2014]. Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position [@ChenSchneider2022; @Karasik2011]. As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014]. Caregivers provide different learning input to infants depending on infants' body position [@KretchKoziol2022; @WestIverson2021; @Karasik2014]. For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects. Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning [@CurrentOpinion].

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a *distal comparison*. Although in the current investigation we focus on classification of body position, this approach can be extended to categorize other aspects of movement, such as locomotion [@AiraksinenRasanen2020] and infant-caregiver contact [@YaoPlotz2019]. We provide openly shared data, materials, and analysis code to facilitate other researchers' adopting this method and applying it to populations and behaviors (see Data Sharing and Transparency below).

@@

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are *mobile*. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is *unobtrusive*. Participants' reactivity to observation, such as from a video camera, may influence behavior. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA&reg; device [@Bergelson2019Input]. Third and fourth, recordings capture *real-time data* over a *full day*. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022; @MalachowskiSalo2023]. Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018]. "Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021]. Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

@@

Third, although the goal was to collect fully naturalistic behavior, we opted to restrict recording to times when participants were in the home. Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands. However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions. Future work should aim to collect data in these scenarios so that body position can be captured in or out of the home.

>>>>

From moment to moment, infants' movements facilitate and constrain how they can interact with their surroundings. Changes in *body position*---whether infants are supine on their backs, prone on their bellies, sitting, upright, or held by a caregiver---have immediate consequences for vision, object exploration, and social interaction. When sitting and upright, infants have a better view of faces and distant objects compared to their view while in a prone position [@Freeplay; @LuoFranchak2020; @CWW]. Learning to sit facilitates object exploration [@SoskaAdolph2014]. Upright walking changes how infants and caregivers interact; walking infants move farther away from caregivers and share toys in different ways compared to infants crawling in a prone position [@ChenSchneider2022; @Karasik2011]. As infants grow older and acquire new abilities, such as independent sitting and walking, they spend more time sitting and upright and less time held, supine, and prone [@Survey; @Thurman2017; @Freeplay; @AdolphCTL2014]. Caregivers provide different learning input to infants depending on infants' body position [@KretchKoziol2022; @WestIverson2021; @Karasik2014]. For example, infants' walking movements spur caregivers to talk about actions rather than just describe objects. Thus, characterizing individual differences in the day-to-day accumulation of body position experiences informs developmental theory by revealing differential opportunities for learning [@CurrentOpinion].

In this paper, we present an inertial sensing method to classify infant body position from moment to moment across an entire day, and validate its accuracy using over 100 hours of video recorded across 34 in-home data collection sessions. Our method takes inspiration from a more mature technology: Long-form audio recordings of infants' language experiences using wearable audio recorders. We begin by describing the impact of long-form audio recordings on research in developmental psychology, and identify the key features that should be replicated in long-form recordings of motor behavior. Next, we review the current state-of-the-art in measuring infant motor behavior---video and survey data---and their limitations in capturing real-time, full-day behavior. Finally, we discuss the advantages of using inertial sensing to classify motor behavior. Despite promising past results in brief, supervised sessions [@FranchakScott2021; @AiraksinenRasanen2020; @AiraksinenGallen2022], the current investigation takes a needed step forward by testing accuracy over long, unsupervised recordings in what we term a *distal comparison*.

@@

We identified five key features of long-form audio methods that should be replicated in analogous studies of motor behavior. First, wearable audio recorders are *mobile*. Measurement is not limited to a particular room because the recording device travels with the participant. Data are recorded to onboard device memory, so participants do not need to be in range of a receiver. Second, wearable audio recording is more *unobtrusive*. Participants' reactivity to observation, such as from a video camera, may influence behavior more compared with a sensor that records only motion or audio data. For example, caregivers spoke more frequently to infants during a video-recorded portion of a home recording compared with audio-only segments captured by a LENA&reg; device [@Bergelson2019Input]. Third and fourth, recordings capture *real-time data* over a *full day*. The ability to record real-time data is vital for making inferences about processes that happen on the timescale of minutes or even seconds within an individual as opposed to comparisons of aggregated data between infants. Synchronizing real-time data to other data streams helps to reveal sources of variability within an individual [e.g., @WassPhillips2022; @MalachowskiSalo2023]. Full-day recordings are essential for capturing experiences across the heterogeneity of daily routines that moderate behavior (e.g., play, feeding, errands) [@KadookaCaufield2021; @Tamis-LeMondaCustode2018]. "Burstiness" of behavior means that long recordings are needed to capture clusters of events amid long periods in which they may be absent [@BarbaroFausey2022; @WarlaumontSobowale2021]. Fifth, *automatic classification* means that the approach can scale to analyze large numbers of participants over long recordings without the bottleneck of manual annotation/transcription.

@@

Third, although the goal was to collect fully naturalistic behavior, our recording protocol led to some changes that might have affected (or missed) some behaviors. We opted to restrict recording to times when participants were in the home. Inertial sensors can travel with participants, and there is no doubt that infants could have worn the sensors out of the home on errands. However, we do not yet have training data to detect periods of transportation (e.g., moving in a car or in a stroller) that might add noise and/or lead to misclassification of body positions. It is also unknown whether participants might have behaved differently had they not worn the sensors. The added bulk in the garment from the sensors (including the LENA&reg; recorder) might have made some positions more uncomfortable, such as lying prone with the LENA&reg; recorder worn on the chest.
